---
title: "Technical Report"
author: "Hazel Ethier and Eva Gerstle"
date: "5/8/2018"
output: pdf_document
---

\newpage

```{r include=FALSE, message=FALSE}
library(tidyverse)
library(ipumsr)
library(mosaic)
library(Stat2Data)
library(dplyr)
library(ggplot2)
library(lmtest)
setwd("~/Desktop")

project_ddi <- read_ipums_ddi("usa_00005.xml")
project_data <- read_ipums_micro(project_ddi, verbose = FALSE)

```

```{r}
project_data <- project_data %>%
  mutate(
    Northeast=if_else(REGION == 11 | REGION == 12, 1 ,0),
    Midwest=if_else(REGION == 21 | REGION == 22, 1, 0),
    South=if_else(REGION == 31 | REGION == 32 | REGION == 33,1,0),
    West=if_else(REGION == 41 | REGION == 42,1,0)
  ) %>%
  mutate(
    white=if_else(RACE == 1,1,0),
    black =if_else(RACE == 2,1,0),
    nativeamerican=if_else(RACE==3,1,0),
    asian=if_else(RACE==4 | RACE==5 | RACE==6,1,0),
    otherrace=if_else(RACE==7,1,0),
    twoormore=if_else(RACE==8 | RACE == 9,1,0)
  ) %>%
  mutate(
    CINETHH= as.factor(if_else(CINETHH == 1 | CINETHH == 2, 1,0))
  ) %>%
  mutate(
    FARM= as.factor(if_else(FARM == 2, 1,0))
  ) %>%
  mutate(
    SSMC= as.factor(if_else(SSMC == 1 | SSMC == 2, 1,0))
  ) %>%
  mutate(
    MULTGEN= as.factor(if_else(MULTGEN == 3, 1,0))
  ) %>%
  select(PERNUM, SERIAL, Northeast, Midwest, South, West, AGE, RACE, SSMC, FARM, CINETHH, MULTGEN, HHINCOME, NCHLT5, white, black, nativeamerican, asian, otherrace, twoormore, NUMPREC)
str(project_data, give.attr = FALSE)
```
#Variable Analysis
```{r}
tally(~Northeast, data=project_data)
tally(~Midwest, data=project_data)
tally(~South, data=project_data)
tally(~West, data=project_data)
tally(~RACE, data=project_data)
tally(~SSMC, data=project_data)
tally(~FARM, data=project_data)
tally(~CINETHH, data=project_data)
tally(~MULTGEN, data=project_data)
favstats(~HHINCOME, data=project_data)
favstats(~NCHLT5, data=project_data)
```

RACE is a categorical variable that refers to the race of the person sampled. 2412703 people sampled are white (code 1), 323250 people are black (code 2), 35633 are Native American (code 3), 42220 are Chinese (code 4), 8138 are Japanese (code 5), 118510 are other Asian (code 6), 121302 are a different race that was no specified (code 7), 83415 are two major races (code 8), and 11316 are three major races (code 9). We also created indicator variables for each of these categories that we plan to use in our analysis of the data. 

SSMC is a binary variable that reports whether the head of household and spouse are a same-sex married couple. 3142297 of our observations were not a same-sex married couple (code 0) while 14190 were in a same-sex marriage.

FARM is a binary varibale that records whether or not the household in question lives on a farm. 3119081 households are not located on a farm (code 0) while 37406 households are on a farm (code 1).

MULTGEN is a binary variable that indicates whether there are 3 or more generations living in the same household. 2913283 households have only one or two generations living in the same household (code 0) while 243204 households have three or more generations living in the same household (code 1).

HHINCOME is a quantitative variavle that gives the total household income. Our total household income variable is a very right skewed variable, with a minimum value of -21,500 dollars, and a mean of 559,783 dollars, while the median value is 74,020 dollars. Our mean and median are quite different indicating the skew. The minimum value is notable due to its negativeness, but we have concluded that some people have negative household income due to debt.  

NCHLT5 is a quantitative variable that gives the number of children age 4 and under residing in the household. Number of Own Children under 5 is also a very right skewed variable, with a mode at 0. This makes logical sense, only a few people have children under the age of 5. The median is 0 while the mean is .0883. The max value is 8, which is quite a lot of children under 5!

CINETHH is a binary variable that indicates whether the household has internet access.Internet access reports whether the household has access to the internet. In 2016, 2,687,693 households have internet access or 85% of households surveyed.

Northeast, Midwest, South, and West are all binary variables that indicate the four United States regions as divided by the census. The survey finds 558073 people in the Northeast or 17.6%. The survey finds 21.5% of people in the Midwest. The South has 37.3% of people, while the West has 23.4% of people. 

The United States Census divides the country into 4 regions: Northeast (Connecticut, Maine, Massachusetts, New Hampshire, Rhode Island, Vermont, New Jersey, New York, Pennsylvania), Midwest (Illinois, Indiana, Michigan, Ohio, Wisconsin, Iowa, Kansas, Minnesota, Missouri, Nebraska, North Dakota, South Dakota), South (Delaware, District of Columbia, Florida, Georgia, Maryland, North Carolina, South Carolina, Virginia, West Virginia, Alabama, Kentucky, Mississippi, Tennessee, Arkansas, Louisiana, Oklahoma, Texas), and West (Arizona, Colorado, Idaho, Montana, Nevada, New Mexico, Utah, Wyoming, Alaska, California, Hawaii, Oregon, Washington).

#Results

##Primary Question 

###Assumptions and Conditions:



\textbf{Randomization:}

We got our data from the American Community Survey on IPUMS USA. This survey is a 1-in-100 national random sample of the population. Because the sampling method is randomized, this condition is met. 

\textbf{Independence:}

Since our data is from the American Community Survey on IPUMS USA which uses a random sampling method and has a large sample size, we do not have any reason to think that whether one household is intergenerational would impact if another household is intergenerational. 

\textbf{Linearity:}

Our data consists of a binary response variable of whether a household is intergenerational and a categorical explanatory variable that we made into three indicator variables. Since we can connect a straight line between any two binary variables, our data is linear. 

```{r}
primarymodel <- glm(MULTGEN ~ West + Midwest + South + AGE + HHINCOME + NUMPREC + NCHLT5 + NUMPREC*NCHLT5 + FARM + black + nativeamerican + asian +otherrace + twoormore , family=binomial, data = project_data)
summary(primarymodel)
```
```{r}
exp(coef(primarymodel))
```

```{r}
#Likelihood Test for Interaction Between Number of Children Under 5 and Number of People in the House

model2<- glm(MULTGEN ~ West + Midwest + South + AGE + HHINCOME + NUMPREC + NCHLT5 + FARM + black + nativeamerican + asian +otherrace + twoormore , family=binomial, data = project_data)

lrtest(model2, primarymodel)
```
```{r}
#Nested Likelihood Test for Region
primarymodel2<-glm(MULTGEN ~ AGE + HHINCOME + NUMPREC + NCHLT5 + NUMPREC*NCHLT5 + FARM + black + nativeamerican + asian +otherrace + twoormore , family=binomial, data = project_data)
lrtest(primarymodel2, primarymodel)
```

##Secondary Question

\textbf{Randomization:}

We got our data from the American Community Survey on IPUMS USA. This survey is a 1-in-100 national random sample of the population. Because the sampling method is randomized, this condition is met. 

\textbf{Independence:}

Since our data is from the American Community Survey on IPUMS USA which uses a random sampling method and has a large sample size, we do not have any reason to think that whether one household is intergenerational would impact if another household is intergenerational. 

\textbf{Linearity:}

Insert graphic here of income vs. intergenerational 

<<<<<<< HEAD
```{r}
#Likelihood Test for Income
modelincome<-glm(MULTGEN ~ West + Midwest + South + AGE + NUMPREC + NCHLT5 + NUMPREC*NCHLT5 + FARM + black + nativeamerican + asian +otherrace + twoormore , family=binomial, data = project_data)
lrtest(modelincome, primarymodel)

```

##Tertiary Question

```{r, echo=FALSE}
tertiarydata <- project_data %>%
  filter(MULTGEN == 1)


justincomeplot <- ggplot(tertiarydata, aes(x=NUMPREC, y= HHINCOME)) + geom_jitter(alpha =.1, color ="blue") 
justincomeplot

m1 <- lm(HHINCOME~NUMPREC, data=tertiarydata)
summary(m1)

modeltertiary <- lm(HHINCOME~NUMPREC + Midwest*NUMPREC + South*NUMPREC + West*NUMPREC + AGE + FARM + NCHLT5 + black*NUMPREC + nativeamerican*NUMPREC + asian*NUMPREC + otherrace*NUMPREC + twoormore*NUMPREC + CINETHH + SSMC, data=tertiarydata)

modeltertiarysmall <- lm(HHINCOME~NUMPREC + Midwest*NUMPREC + South*NUMPREC + West*NUMPREC, data=tertiarydata)
anova(modeltertiarysmall, modeltertiary)
summary(modeltertiarysmall)
summary(modeltertiary)
```
```{r}
tally(~MULTGEN, data=project_data)
```


The United States Census divides the country into 4 regions: Northeast (Connecticut, Maine, Massachusetts, New Hampshire, Rhode Island, Vermont, New Jersey, New York, Pennsylvania), Midwest (Illinois, Indiana, Michigan, Ohio, Wisconsin, Iowa, Kansas, Minnesota, Missouri, Nebraska, North Dakota, South Dakota), South (Delaware, District of Columbia, Florida, Georgia, Maryland, North Carolina, South Carolina, Virginia, West Virginia, Alabama, Kentucky, Mississippi, Tennessee, Arkansas, Louisiana, Oklahoma, Texas), and West (Arizona, Colorado, Idaho, Montana, Nevada, New Mexico, Utah, Wyoming, Alaska, California, Hawaii, Oregon, Washington). [^2]

```{r, message=FALSE}
plot(modeltertiary, which=1)
plot(modeltertiary, which=2)
```


#Discussion

\textbf{Linearity:} There is mostly random scatters in the fitted residuals plot (and no obvious curvature) so the data appears linear enough.

\textbf{Independence:} Since our data is from the American Community Survey on IPUMS USA which uses a random sampling method and has a large sample size, we do not have any reason to think that the household income of one household affects the household income of another household.

\textbf{Normality:} Unfortunately, our data does not meet the normality condition, as the points curve off of the Normal Q-Q plot.

\textbf{Zero Mean:} In the fitted residuals plot, the residuals are equally scattered above and below the mean therefore I can conclude that the residuals have a mean of zero. 

\textbf{Equal Variance:}The residuals in the fitted residuals plot seem to have an equal width and therefore a uniform standard deviation. There does not appear to be any thickening.

